---
title: "Sitka_scratch"
author: "Luke Miller"
date: "8/19/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
evalall = FALSE
showcode = FALSE
```

```{r loadPackages}
library(signal)
library(pracma)
library(forecast)

```


```{r importSN05, eval = evalall}
# Load the couple of SN05 raw data files
files = dir('../Data/2020_Sitka_mussel_data/Raw_files/', full.names=TRUE)

for (i in 1:length(files)){
	temp = read.csv(files[i])
	if (i == 1) {
		SN05 = temp
	} else if (i > 1){
		SN05 = rbind(SN05,temp)
	}
}
# Convert time stamps to POSIX
SN05$DateTime = as.POSIXct(SN05$DateTime, tz = 'UTC')
SN05$DateTimeAST = SN05$DateTime
attr(SN05$DateTimeAST,'tzone') = 'etc/GMT+9' # Create Alaska Standard Time version

```





```{r testheartratefilter, echo=showcode , eval=evalall} 
# Some manual test code used to develop the beats-per-minute calculations
# and filtering settings

Fs = 0.1 # 10Hz sampling rate, sampling interval = 0.1 seconds
# Step through 1 minute at a time
mytime = as.POSIXct('2020-09-20 20:15',tz='etc/GMT+9')

t1 = which.min(abs(SN05$DateTimeAST - mytime))
chunklength = 300 # number of samples to use
# Test that we've hit the chosen minute (not just the closest time)
if (difftime(mytime,SN05$DateTimeAST[t1], units = 'secs') == 0){
	# Extract a chunk of samples slightly longer than target length
	temp = SN05[t1:(t1+(chunklength*2)),]
	diffs = diff(temp$startMillis)
	missedReads = which(diffs > 100)
	# Check if there are any sampling gaps (missedReads)
	if (length(missedReads) > 0){
		testgaps = diff(missedReads) 
		if (missedReads[1] > 300){
			# In this case, just grab the first 300 readings
			temp = temp[1:chunklength,]
		} else if (length(which(testgaps>=300)) > 0){
			# check if any of the testgaps values are >=300
			# Get the index in temp that is at the start of the long run
			tempindx = missedReads[which(testgaps >= 300)]
			# If a gap is > 300, grab the sample in that gap
			temp = temp[tempindx:(tempindx+chunklength-1),]
		} else if (length(which(testgaps>=300)) == 0) {
			# In this case there are good data from the last missedRead until
			# the end of the temp dataframe, so just grab the chunk from 
			# the last missedRead
			tempindx = missedReads[length(missedReads)]+1
			temp = temp[tempindx:(tempindx+chunklength-1),]
			#TODO you can hit a situation where you end up in this section
			# when there are gaps spaced such that you can't get a contiguous
			# 300 samples within the 600 sample chunk you grabbed. That 
			# needs to be handled somehow, currently it just fails. 
			
		}
	} else if (length(missedReads) == 0) {
		# Subset down to exactly the chunk length (10Hz sample * 30 secs = 300 samples)
		temp = temp[1:chunklength,]
	}

	# Add milliseconds onto the timestamps
	temp$DateTimeMS = temp$DateTimeAST
	
	diffs = diff(temp$startMillis) / 1000
	temp$DateTimeMS[2:nrow(temp)] = temp$DateTimeMS[1] + cumsum(diffs)
		
	detrendedIR = detrend(temp[,'Sensor8IR'])
	
	# Define a butterworth filter
	# Consider using the bandpass filter rather than just a lowpass filter
	# because of the tendency for the IR heartrate signal to drift up and 
	# down in relation to ambient light, which tends to induce low-frequency
	# shifts that then fool the spectral analysis routines when trying to 
	# identify the dominant frequency
	bf = butter(3,W = c(1/60, 1/10), type = 'pass')  # bandpass filter
#	bf2 = butter(3,W = c(1/120, 1/10), type = 'pass')  # bandpass filter
#	bf = butter(3,W = 0.1, type = 'low')  # 10Hz lowpass filter
	# Apply the filter to the detrended data chunk
	y = filtfilt(bf, x = detrendedIR)
#	y2 = filtfilt(bf2, x = detrendedIR)
	myfft2 = spectrum(y, plot = FALSE)
	# Take the peak frequency from the spectrum, divide by sampling
	# rate to convert to cycles per second
	mypeakfreq = myfft2$freq[which.max(myfft2$spec)] / Fs
	# Calculate amplitude of filtered signal to avoid weak or noisy signals
	amp = range(y)[2] - range(y)[1]
	# Multiply by 60 seconds to get cycles (beats) per minute
	bpm = 60 * mypeakfreq
	# Plot the raw detrended signal
	plot(temp$DateTimeMS, detrendedIR, type = 'l', 
			main = paste0("Amplitude = ",round(amp,1),
					', BPM = ', bpm), 
			las = 1) 
	if (amp > 60){
		lines(temp$DateTimeMS,y, col = 3, lwd = 2) # add the filtered signal
#		lines(temp$DateTimeMS,y2, col = 4, lwd = 2)
	} else if (amp <= 60) {
		lines(temp$DateTimeMS,y, col = 2, lwd = 2) # add the filtered signal	
		warning('Garbage signal')
	}
	# Use function from package 'forecast', returns peak period (not freq)
	forecastPeriod = forecast::findfrequency(y)  
	forecastFreq = 1/forecastPeriod # convert period to frequency
	forecastBPM = forecastFreq * 60 * 10 # convert frequency to beats per 
	# minute, based on the fact that the sampling rate is 10Hz, and there are
	# 60 seconds in a minute
	mtext(side = 3, text = paste0('Forecast BPM: ',round(forecastBPM,1)))
	# Use pracma package to find peaks. Note that at slower heart rates
	# this function tends to find the sub-peaks (akin to a P or T peak in a
	# human ECG trace) rather than just the main peaks (R peaks on a human). 
	# This happens based on what the bandpass filter lets through. 
	res = pracma::findpeaks(x = y, nups = 5, minpeakdistance = 10)
	points(temp$DateTimeMS[res[,2]], y = res[,1], col = 4, pch = 19)
	# Print the estimated bpm from the pracma::findpeaks routine, but 
	# you need to double the number of peaks to get bpm if your time series
	# is only 30 seconds long
	mtext(side = 3, line = 3, text = paste0('pracma bpm: ', nrow(res)*2))
} 


``` 

```{r heartrateFilter,echo=showcode}
# A function to convert raw heartbeat timeseries into estimates of heart rate
# (beats per minute). 

#' @param x A data frame with columns DateTimeAST, startMillis, and sensor data
#' @param t1 The starting index of data to be subset from x
#' @param sensor Column name in x declaring the sensor that you want to process
#' @param chunklength Number of samples to process, default = 300 is 30 seconds at 10Hz
#' @param lowband Frequency for the lower frequency of the bandpass filter
#' @param highband Frequency for the upper frequency of the bandpass filter
#' @param Fs Sampling interval in seconds (default 0.1 = 10Hz)
#' 
#' @return A list containing 2 estimates of the beats per minute, along with
#' values indicating how close the 2 estimates agree and whether the signal may
#' be questionable.

getBPM = function(x,t1 = 1, sensor = 'Sensor1IR',chunklength=300, 
		lowband = 1/60, 
		highband = 1/10, 
		Fs = 0.1)
{
	# Grab a chunk of data that's twice as long as we need
	temp = x[t1:(t1+(chunklength*2)),]
	# Calculate time difference between each sample (milliseconds), should be 100
	diffs = diff(temp$startMillis)
	# Identify any gaps in the data where the interval was >100ms
	missedReads = which(diffs > 100)
# Check if there are any sampling gaps (missedReads)
	if (length(missedReads) > 0){
		testgaps = diff(missedReads) 
		if (missedReads[1] > 300){
			# In this case, just grab the first 300 readings
			temp = temp[1:chunklength,]
		} else if (length(which(testgaps>=300)) > 0){
			# check if any of the testgaps values are >=300
			# Get the index in temp that is at the start of the long run
			tempindx = missedReads[which(testgaps >= 300)]
			# If a gap is > 300, grab the sample in that gap
			temp = temp[tempindx:(tempindx+chunklength-1),]
		} else if (length(which(testgaps>=300)) == 0) {
			# In this case there are good data from the last missedRead until
			# the end of the temp dataframe, so just grab the chunk from 
			# the last missedRead
			tempindx = missedReads[length(missedReads)]+1
			temp = temp[tempindx:(tempindx+chunklength-1),]
			# It's also possible to end up here if you grabbed a chunk that
			# ran off the end of the 600 samples and returned some NAs. That
			# will be handled below
			
		}
	} else if (length(missedReads) == 0) {
		# Subset down to exactly the chunk length (10Hz sample * 30 secs = 300 samples)
		temp = temp[1:chunklength,]
	}

	# Detrend the heartrate readings
	detrendedIR = pracma::detrend(temp[,sensor])
	# Test if there are any NAs in the heart rate values, if there are not
	# then proceed with the filtering and heart rate determination
	if (length(which(is.na(detrendedIR))) == 0) {
		# Define a butterworth filter
	# Consider using the bandpass filter rather than just a lowpass filter
	# because of the tendency for the IR heartrate signal to drift up and 
	# down in relation to ambient light, which tends to induce low-frequency
	# shifts that then fool the spectral analysis routines when trying to 
	# identify the dominant frequency
		bf = signal::butter(3,W = c(lowband, highband), type = 'pass')
		
		# Apply the filter to the detrended data chunk
		y = filtfilt(bf, x = detrendedIR)
		# Calculate the spectrum of the filtered data
		myfft2 = spectrum(y, plot = FALSE)
		# Take the peak frequency from the spectrum, divide by sampling
		# rate to convert to cycles per second
		mypeakfreq = myfft2$freq[which.max(myfft2$spec)] / Fs
		# Multiply by 60 seconds to get cycles (beats) per minute
		BPMfft = 60 * mypeakfreq
		# Calculate amplitude of filtered signal to denote weak or noisy signals
		amp = range(y)[2] - range(y)[1]
		
		## Plot the raw detrended signal
#	plot(temp$DateTimeMS, detrendedIR, type = 'l', 
#			main = paste0("Amplitude = ",round(amp,1),
#					', BPM = ', bpm), 
#			las = 1) 
#	if (amp > 60){
#		lines(temp$DateTimeMS,y, col = 3, lwd = 2) # add the filtered signal
#	} else if (amp <= 60) {
#		lines(temp$DateTimeMS,y, col = 2, lwd = 2) # add the filtered signal	
#		warning('Garbage signal')
#	}
		# Use function from package 'forecast', returns peak period (not freq)
		forecastPeriod = forecast::findfrequency(y)  
		forecastFreq = 1/forecastPeriod # convert period to frequency
		forecastBPM = forecastFreq * 60 * 10 # convert frequency to beats per 
		# minute, based on the fact that the sampling rate is 10Hz, and there are
		# 60 seconds in a minute
		#	mtext(side = 3, text = paste0('Forecast BPM: ',round(forecastBPM,1)))
		# Use pracma package to find peaks. Note that at slower heart rates
		# this function tends to find the sub-peaks (akin to a P or T peak in a
		# human ECG trace) rather than just the main peaks (R peaks on a human). 
		# This happens based on what the bandpass filter lets through. 
#	res = pracma::findpeaks(x = y, nups = 5, minpeakdistance = 10)
#	points(temp$DateTimeMS[res[,2]], y = res[,1], col = 4, pch = 19)
		# Print the estimated bpm from the pracma::findpeaks routine, but 
		# you need to double the number of peaks to get bpm if your time series
		# is only 30 seconds long
#	mtext(side = 3, line = 3, text = paste0('pracma bpm: ', nrow(res)*2))
		
		# Calculate the difference between the estimated heart rates from the 
		# spectrum fft routine and the forecast::findfrequency routine. A 
		# small value indicates good agreement
		BPMagreeQuality = ceiling(abs(BPMfft - forecastBPM))
		# Set a flag for cases where the forecast:findfrequency estimate is 
		# extremely large, which happens when it can't find a clear heart signal
		forecastBPMflag = ifelse(forecastBPM > 60, 'FAIL','OK')
		# Set a flag for cases where the detrended/filtered signal has a very
		# small amplitude, signaling that there may be no good heartbeat signal 
		WeakSignalFlag = ifelse(amp < 60, 'FAIL','OK')
		
		
		resultsList = list(DateTimeAST = temp$DateTimeAST[1],
				Sensor = sensor,
				BPMfft = round(BPMfft,1),
				forecastBPM = round(forecastBPM,1),
				BPMagreeQuality = BPMagreeQuality,
				forecastBPMflag = forecastBPMflag,
				WeakSignalFlag = WeakSignalFlag)
	} else if (length(is.na(detrendedIR)) > 0) {
		# Handle the case where there are NAs in the data chunk that prevent
		# the filtering and fft routines
		resultsList = list(DateTimeAST = temp$DateTimeAST[1],
				Sensor = sensor,
				BPMfft = NA,
				forecastBPM = NA,
				BPMagreeQuality = NA,
				forecastBPMflag = 'FAIL',
				WeakSignalFlag = 'FAIL')
	}
	

	return(resultsList)
}



```


```{r processSN05heartrates,echo=showcode, eval = evalall}
# Go through the different sensors in the raw SN05 data frame and 
# calculate beats per minute estimates. The resulting data will be
# written to a csv file for faster reloading in the future. This 
# routine takes about 10-15 minutes for all 8 sensors to process. 

# For SN05 these mussels were submerged in the pools: 1, 2, 3, 6
# For SN05 these mussels were out on the edge of the pool: 4, 5, 7, 8

# Step through 1 minute at a time
mytime = as.POSIXct('2020-09-20 15:00',tz='etc/GMT+9')

timesteps = seq(SN05$DateTimeAST[1],SN05$DateTimeAST[nrow(SN05)],by = 60)

# Sensor1
#for (i in 1:16){  # for testing
for (i in 1:length(timesteps)){
	t1 = which.min(abs(SN05$DateTimeAST - timesteps[i]))
	# Test that we've hit the chosen minute (not just the closest time)
	if (difftime(timesteps[i],SN05$DateTimeAST[t1], units = 'secs') == 0){
		
		# Sensor 1
		res = getBPM(SN05,t1 = t1, sensor = 'Sensor1IR')
		if (i == 1) {			
			SensorResult = as.data.frame(res)
			SensorResult[2:length(timesteps),] = NA
		} else {
			SensorResult[i,] = as.data.frame(res)
		}
	} else {
		SensorResult$DateTimeAST[i] = timesteps[i]
	}
}
SN05S1 = SensorResult
SN05S1$Treatment = 'Submerged'

# Sensor2
for (i in 1:length(timesteps)){
#for (i in 1:16){
	t1 = which.min(abs(SN05$DateTimeAST - timesteps[i]))
	# Test that we've hit the chosen minute (not just the closest time)
	if (difftime(timesteps[i],SN05$DateTimeAST[t1], units = 'secs') == 0){
		
		# Sensor 1
		res = getBPM(SN05,t1 = t1, sensor = 'Sensor2IR')
		if (i == 1) {			
			SensorResult = as.data.frame(res)
			SensorResult[2:length(timesteps),] = NA
		} else {
			SensorResult[i,] = as.data.frame(res)
		}
	} else {
		SensorResult$DateTimeAST[i] = timesteps[i]
	}
}
SN05S2 = SensorResult
SN05S2$Treatment = 'Submerged'

# Sensor3
for (i in 1:length(timesteps)){
#for (i in 1:16){
	t1 = which.min(abs(SN05$DateTimeAST - timesteps[i]))
	# Test that we've hit the chosen minute (not just the closest time)
	if (difftime(timesteps[i],SN05$DateTimeAST[t1], units = 'secs') == 0){
		
		# Sensor 1
		res = getBPM(SN05,t1 = t1, sensor = 'Sensor3IR')
		if (i == 1) {			
			SensorResult = as.data.frame(res)
			SensorResult[2:length(timesteps),] = NA
		} else {
			SensorResult[i,] = as.data.frame(res)
		}
	} else {
		SensorResult$DateTimeAST[i] = timesteps[i]
	}
}
SN05S3 = SensorResult
SN05S3$Treatment = 'Submerged'

# Sensor4
for (i in 1:length(timesteps)){
#for (i in 1:16){
	t1 = which.min(abs(SN05$DateTimeAST - timesteps[i]))
	# Test that we've hit the chosen minute (not just the closest time)
	if (difftime(timesteps[i],SN05$DateTimeAST[t1], units = 'secs') == 0){
		
		# Sensor 1
		res = getBPM(SN05,t1 = t1, sensor = 'Sensor4IR')
		if (i == 1) {			
			SensorResult = as.data.frame(res)
			SensorResult[2:length(timesteps),] = NA
		} else {
			SensorResult[i,] = as.data.frame(res)
		}
	} else {
		SensorResult$DateTimeAST[i] = timesteps[i]
	}
}
SN05S4 = SensorResult
SN05S4$Treatment = 'Emersed'

# Sensor5
for (i in 1:length(timesteps)){
#for (i in 1:16){
	t1 = which.min(abs(SN05$DateTimeAST - timesteps[i]))
	# Test that we've hit the chosen minute (not just the closest time)
	if (difftime(timesteps[i],SN05$DateTimeAST[t1], units = 'secs') == 0){
		
		# Sensor 1
		res = getBPM(SN05,t1 = t1, sensor = 'Sensor5IR')
		if (i == 1) {			
			SensorResult = as.data.frame(res)
			SensorResult[2:length(timesteps),] = NA
		} else {
			SensorResult[i,] = as.data.frame(res)
		}
	} else {
		SensorResult$DateTimeAST[i] = timesteps[i]
	}
}
SN05S5 = SensorResult
SN05S5$Treatment = 'Emersed'

#Sensor 6
for (i in 1:length(timesteps)){
#for (i in 1:16){
	t1 = which.min(abs(SN05$DateTimeAST - timesteps[i]))
	# Test that we've hit the chosen minute (not just the closest time)
	if (difftime(timesteps[i],SN05$DateTimeAST[t1], units = 'secs') == 0){
		
		# Sensor 1
		res = getBPM(SN05,t1 = t1, sensor = 'Sensor6IR')
		if (i == 1) {			
			SensorResult = as.data.frame(res)
			SensorResult[2:length(timesteps),] = NA
		} else {
			SensorResult[i,] = as.data.frame(res)
		}
	} else {
		SensorResult$DateTimeAST[i] = timesteps[i]
	}
}
SN05S6 = SensorResult
SN05S6$Treatment = 'Submerged'

# Sensor7

for (i in 1:length(timesteps)){
#for (i in 1:16){
	t1 = which.min(abs(SN05$DateTimeAST - timesteps[i]))
	# Test that we've hit the chosen minute (not just the closest time)
	if (difftime(timesteps[i],SN05$DateTimeAST[t1], units = 'secs') == 0){
		
		# Sensor 1
		res = getBPM(SN05,t1 = t1, sensor = 'Sensor7IR')
		if (i == 1) {			
			SensorResult = as.data.frame(res)
			SensorResult[2:length(timesteps),] = NA
		} else {
			SensorResult[i,] = as.data.frame(res)
		}
	} else {
		SensorResult$DateTimeAST[i] = timesteps[i]
	}
}
SN05S7 = SensorResult
SN05S7$Treatment = 'Emersed'

# Sensor 8
# There are problems on this one because most of the later minutes have
# >100ms gaps in them that don't allow a contiguous 300 samples to be extracted
# and you end up with NAs in the dataset, which leads to NAs in the output
for (i in 1:length(timesteps)){
#for (i in 1:16){
	t1 = which.min(abs(SN05$DateTimeAST - timesteps[i]))
	# Test that we've hit the chosen minute (not just the closest time)
	if (difftime(timesteps[i],SN05$DateTimeAST[t1], units = 'secs') == 0){
		
		# Sensor 1
		res = getBPM(SN05,t1 = t1, sensor = 'Sensor8IR')
		if (i == 1) {			
			SensorResult = as.data.frame(res)
			SensorResult[2:length(timesteps),] = NA
		} else {
			SensorResult[i,] = as.data.frame(res)
		}
	} else {
		SensorResult$DateTimeAST[i] = timesteps[i]
	}
}
SN05S8 = SensorResult
SN05S8$Treatment = 'Emersed'

SN05bpm = rbind(SN05S1,SN05S2,SN05S3,SN05S4,SN05S5,SN05S6,SN05S7,SN05S8)

write.csv(SN05bpm,
		file = '../Data/2020_Sitka_mussel_data/Sitka_SN05_bpm_reprocessed_20200920.csv',
		row.names=FALSE)
```

```{r openSN05,echo=showcode}
# Open the re-processed beats per minute estimates from board SN05
SN05bpm = read.csv('../Data/2020_Sitka_mussel_data/Sitka_SN05_bpm_reprocessed_20200920.csv')
SN05bpm$DateTimeAST = as.POSIXct(SN05bpm$DateTimeAST, tz = 'etc/GMT+9')
SN05bpm$Sensor = factor(SN05bpm$Sensor)
SN05bpm$forecastBPMflag = factor(SN05bpm$forecastBPMflag)
SN05bpm$WeakSignalFlag = factor(SN05bpm$WeakSignalFlag)
SN05bpm$Treatment = factor(SN05bpm$Treatment)

# For SN05 these mussels were submerged in the pools: 1, 2, 3, 6
# For SN05 these mussels were out on the edge of the pool: 4, 5, 7, 8


```

```{r subsetSN05,echo=showcode}

SN05S1 = SN05bpm[SN05bpm$Sensor == 'Sensor1IR',]	# submerged
SN05S2 = SN05bpm[SN05bpm$Sensor == 'Sensor2IR',]	# submerged
SN05S3 = SN05bpm[SN05bpm$Sensor == 'Sensor3IR',]	# submerged
SN05S4 = SN05bpm[SN05bpm$Sensor == 'Sensor4IR',]	# emersed
SN05S5 = SN05bpm[SN05bpm$Sensor == 'Sensor5IR',]	# emersed
SN05S6 = SN05bpm[SN05bpm$Sensor == 'Sensor6IR',]	# submerged
SN05S7 = SN05bpm[SN05bpm$Sensor == 'Sensor7IR',]	# emersed
SN05S8 = SN05bpm[SN05bpm$Sensor == 'Sensor8IR',]  	# emersed

# Remove questional bpm estimates based on the flag values
SN05S1[which(SN05S1$WeakSignalFlag == 'FAIL'),c('BPMfft','forecastBPM')] = NA
SN05S1[which(SN05S1$forecastBPMflag == 'FAIL'),c('forecastBPM')] = NA

SN05S2[which(SN05S2$WeakSignalFlag == 'FAIL'),c('BPMfft','forecastBPM')] = NA
SN05S2[which(SN05S2$forecastBPMflag == 'FAIL'),c('forecastBPM')] = NA

SN05S3[which(SN05S3$WeakSignalFlag == 'FAIL'),c('BPMfft','forecastBPM')] = NA
SN05S3[which(SN05S3$forecastBPMflag == 'FAIL'),c('forecastBPM')] = NA

SN05S4[which(SN05S4$WeakSignalFlag == 'FAIL'),c('BPMfft','forecastBPM')] = NA
SN05S4[which(SN05S4$forecastBPMflag == 'FAIL'),c('forecastBPM')] = NA

SN05S5[which(SN05S5$WeakSignalFlag == 'FAIL'),c('BPMfft','forecastBPM')] = NA
SN05S5[which(SN05S5$forecastBPMflag == 'FAIL'),c('forecastBPM')] = NA

SN05S6[which(SN05S6$WeakSignalFlag == 'FAIL'),c('BPMfft','forecastBPM')] = NA
SN05S6[which(SN05S6$forecastBPMflag == 'FAIL'),c('forecastBPM')] = NA

SN05S7[which(SN05S7$WeakSignalFlag == 'FAIL'),c('BPMfft','forecastBPM')] = NA
SN05S7[which(SN05S7$forecastBPMflag == 'FAIL'),c('forecastBPM')] = NA

SN05S8[which(SN05S8$WeakSignalFlag == 'FAIL'),c('BPMfft','forecastBPM')] = NA
SN05S8[which(SN05S8$forecastBPMflag == 'FAIL'),c('forecastBPM')] = NA

```

```{r importTides,echo=showcode}
tides = read.csv(file = '../Data/2020_Sitka_mussel_data/Sitka_tides_20200919-20200923.csv')


tides$DateTimeUTC = paste(tides$Date, tides$Time..GMT.)
tides$DateTimeUTC = as.POSIXct(tides$DateTimeUTC, format = '%Y/%m/%d %H:%M',
		tz = 'UTC')
tides$DateTimeAST = tides$DateTimeUTC
attr(tides$DateTimeAST, 'tzone') = 'etc/GMT+9'
tides$HighLow = NA
height = 4.1
tides$HighLow[which(tides$Verified..ft. > height)] = 'High'
tides$HighLow[which(tides$Verified..ft. <= height)] = 'Low'
tides$HighLow = factor(tides$HighLow)
```


```{r PlotFunctions}
# Function to plot a heartrate dataset, plotting both the actual points
# and a loess smoother
loessFunc = function(x,y, col = 1, lwd = 2, pch = '.', cex = 1, span = 0.75){
	points(x, y, col = col, pch = pch, cex = cex)
	lo = loess(y~as.numeric(x), na.action = 'na.omit', span = span)
	predy = predict(lo, newdata = as.numeric(x))
	lines(x, predy, col = col, lwd = lwd)
}

TideBoxes = function(Exposure,DateTime, cols = c('lightblue','white')){
	runs = rle(as.numeric(Exposure))
	runs$lengths = c(1,runs$lengths)
	
	indsright = numeric(0)
	for (i in 1:length(runs$lengths)){
		indsright = c(indsright,sum(runs$lengths[1:i])) 
	}
	indsleft = c(1,indsright[1:(length(indsright)-1)])
	
	plotdims = par()$usr
	
	rect(xleft = DateTime[indsleft], 
			ybottom = rep(plotdims[3],length(indsleft)),
			xright = DateTime[indsright],
			ytop = rep(plotdims[4], length(indsleft)),
			col = cols, border = NA)
	box()
}

```

TODO: I still think there's a discrepancy between the mussel time values and
the real world. It's possible that SN05 board wasn't set to UTC time zone, but
instead to maybe Pacific Daylight Time (UTC-7), Alaska Daylight Time (UTC-8),
or Alaska Standard Time (UTC-9). If so, my assumption that it was UTC, and 
then subsequently producing a UTC-9 AKST version of the timestamp would be 
throwing off the data by several hours.  

```{r plotSN05BPM1, echo=showcode}
png(filename = './figs/SN05_heartrates_20200919-20200921.png',
		width = 5, height = 5, units = 'in',res=300)
par(mar = c(5,5.5,1,2))
mypch = '.'
mycex = 1.5
myspan = 0.1
plot(BPMfft~DateTimeAST,
		data = subset(SN05S1, BPMagreeQuality < 2, na.action='na.omit'), 
		col = 1, pch = mypch, type = 'n', las = 1, cex = mycex,
		ylim = c(0,35), 
		xlab = 'Sep 19-22, 2020',
		ylab = 'Heart rate (bpm)',
		xaxs = 'i')
# Trying to fix the time zone offset issue
TideBoxes(tides$HighLow,(tides$DateTimeAST + (11*3600)))
t1 = subset(SN05S1, BPMagreeQuality < 2, na.action = 'na.omit')
loessFunc(t1$DateTimeAST, t1$forecastBPM, col = 4, lwd = 2, cex = mycex, span = myspan)
t1 = subset(SN05S2, BPMagreeQuality < 2, na.action = 'na.omit')
loessFunc(t1$DateTimeAST, t1$forecastBPM, col = 4, lwd = 2, cex = mycex, span = myspan)
t1 = subset(SN05S3, BPMagreeQuality < 2, na.action = 'na.omit')
loessFunc(t1$DateTimeAST, t1$forecastBPM, col = 4, lwd = 2, cex = mycex, span = myspan)
t1 = subset(SN05S6, BPMagreeQuality < 2, na.action = 'na.omit')
loessFunc(t1$DateTimeAST, t1$forecastBPM, col = 4, lwd = 2, cex = mycex, span = myspan)
# Emersed mussels
t1 = subset(SN05S4, BPMagreeQuality < 2, na.action = 'na.omit')
loessFunc(t1$DateTimeAST, t1$forecastBPM, col = 2, lwd = 2, cex = mycex, span = myspan)
t1 = subset(SN05S5, BPMagreeQuality < 2, na.action = 'na.omit')
loessFunc(t1$DateTimeAST, t1$forecastBPM, col = 2, lwd = 2, cex = mycex, span = myspan)
t1 = subset(SN05S7, BPMagreeQuality < 2, na.action = 'na.omit')
loessFunc(t1$DateTimeAST, t1$forecastBPM, col = 2, lwd = 2, cex = mycex, span = myspan)
t1 = subset(SN05S8, BPMagreeQuality < 2, na.action = 'na.omit')
loessFunc(t1$DateTimeAST, t1$forecastBPM, col = 2, lwd = 2, cex = mycex, span = myspan)

#legend('topright', legend = c('In-pool mussels','Emersed mussels'), 
#		col = c(4,2), lwd = 2, bg = 'white')
legend('topright', legend = c('In-pool mussels','Emersed mussels','High Tide'), 
		col = c(4,2, 'lightblue'), lwd = c(2,2,8), bg = 'white', cex = 0.8)
dev.off()
```


<!-- ```{r loadFiles} -->
<!-- dat = read.csv("../Data/20200821_1851_01_SN03_IR.csv") -->
<!-- temps = read.csv("../Data/20200821_1851_01_SN03_TEMP.csv") -->
<!--  -->
<!-- # Field data from Gabby, September 2020 -->
<!-- # Note that this file has a lot of duplicate entries in it, so treat it carefully -->
<!-- sitka = read.csv("../Data/2020_Sitka_mussel_data/Sitka2020heartrateenvirodata.csv") -->
<!-- sitka$DateTimeAST = as.POSIXct(sitka$DateTimeAST, tz = 'etc/GMT+9') -->
<!-- sitka$SN_S = factor(sitka$SN_S) -->
<!-- sitka$SN = factor(sitka$SN) -->
<!-- sitka$Exposure = factor(sitka$Exposure) -->
<!-- sitka$Treatment = factor(sitka$Treatment) -->
<!--  -->
<!-- mytime = as.POSIXct('2020-09-21', tz = 'etc/GMT+9') -->
<!-- sitka2 = sitka[sitka$DateTimeAST > mytime,] -->
<!-- ``` -->

<!-- ```{r extractMussels} -->
<!--  -->
<!-- #levels(sitka$SN_S) -->
<!-- # Subset the different mussels  -->
<!-- SN03S1 = sitka2[which(sitka2$SN_S == 'SN03Sensor1'),] # inside -->
<!-- SN03S4 = sitka2[which(sitka2$SN_S == 'SN03Sensor4'),] # 	outside -->
<!-- SN03S8 = sitka2[which(sitka2$SN_S == 'SN03Sensor8'),] # inside -->
<!-- SN04S4 = sitka2[which(sitka2$SN_S == 'SN04Sensor4'),] # 	outside -->
<!-- SN04S4 = SN04S4[1440:nrow(SN04S4),] # Remove duplicate stretch -->
<!-- SN04S6 = sitka2[which(sitka2$SN_S == 'SN04Sensor6'),] # inside -->
<!-- SN04S6 = SN04S6[1440:nrow(SN04S6),]  # Remove duplicate stretch -->
<!-- SN04S8 = sitka2[which(sitka2$SN_S == 'SN04Sensor8'),] # 	outside -->
<!-- SN04S8 = SN04S8[1440:nrow(SN04S8),]  # Remove duplicate stretch -->
<!-- SN05S1 = sitka2[which(sitka2$SN_S == 'SN05Sensor1'),] # inside -->
<!-- SN05S2 = sitka2[which(sitka2$SN_S == 'SN05Sensor2'),] # inside -->
<!-- SN05S3 = sitka2[which(sitka2$SN_S == 'SN05Sensor3'),] # inside -->
<!-- SN05S6 = sitka2[which(sitka2$SN_S == 'SN05Sensor6'),] # inside -->
<!-- SN05S7 = sitka2[which(sitka2$SN_S == 'SN05Sensor7'),] # 	outside -->
<!-- SN05S7 = SN05S7[1440:4318,]  # Remove duplicate stretch -->
<!-- ``` -->

<!-- ```{r} -->
<!-- par(mfrow = c(2,1)) -->
<!-- mypch = '.' -->
<!-- mycex = 1.5 -->
<!-- ylims = c(10,50) -->
<!-- # Inside pool mussels -->
<!-- plot(SN03S1$DateTimeAST, SN03S1$heartrate, type = 'p', las = 1, -->
<!-- 		xlab = 'Time AKST', ylab = 'Heart rate (bpm)', -->
<!-- 		main = "Inside pool", pch = mypch, ylim = ylims) -->
<!-- rect(xleft=par()$usr[1],ybottom=par()$usr[3],xright=par()$usr[2], -->
<!-- 		ytop=par()$usr[4], col = 'grey80') -->
<!-- box() -->
<!-- points(SN03S1$DateTimeAST, SN03S1$heartrate, col = 1, pch = mypch, cex = mycex) -->
<!-- points(SN03S8$DateTimeAST, SN03S8$heartrate, col = 2, pch = mypch, cex = mycex) -->
<!-- points(SN04S6$DateTimeAST, SN04S6$heartrate, col = 3, pch = mypch, cex = mycex) -->
<!-- points(SN05S1$DateTimeAST, SN05S1$heartrate, col = 4, pch = mypch, cex = mycex) -->
<!-- points(SN05S2$DateTimeAST, SN05S2$heartrate, col = 5, pch = mypch, cex = mycex) -->
<!-- points(SN05S3$DateTimeAST, SN05S3$heartrate, col = 6, pch = mypch, cex = mycex) -->
<!-- points(SN05S6$DateTimeAST, SN05S6$heartrate, col = 7, pch = mypch, cex = mycex) -->
<!--  -->
<!-- # Outside pools -->
<!-- plot(SN03S4$DateTimeAST, SN03S4$heartrate, type = 'p', las = 1, -->
<!-- 		xlab = 'Time AKST', ylab = 'Heart rate (bpm)', -->
<!-- 		main = 'Outside pool', pch = mypch, ylim = ylims) -->
<!-- rect(xleft=par()$usr[1],ybottom=par()$usr[3],xright=par()$usr[2], -->
<!-- 		ytop=par()$usr[4], col = 'grey80') -->
<!-- box() -->
<!-- points(SN03S4$DateTimeAST, SN03S4$heartrate, col = 1, pch = mypch, cex = mycex) -->
<!-- points(SN04S4$DateTimeAST, SN04S4$heartrate, col = 2, pch = mypch, cex = mycex) -->
<!-- points(SN04S8$DateTimeAST, SN04S8$heartrate, col = 3, pch = mypch, cex = mycex) -->
<!-- points(SN05S7$DateTimeAST, SN05S7$heartrate, col = 4, pch = mypch, cex = mycex) -->
<!--  -->
<!--  -->
<!-- ``` -->




<!-- ```{r smoothedHeartRatePlot,fig.width=8,fig.height=5,dpi=600} -->
<!-- mypch = '.' -->
<!-- mycex = 1.5 -->
<!-- ylims = c(10,50) -->
<!-- myspan = 0.25 -->
<!-- plot(SN03S4$DateTimeAST, SN03S4$heartrate, type = 'p', las = 1, -->
<!-- 		xlab = 'Time, AKST', ylab = 'Heart rate (bpm)', -->
<!-- 		main = '', pch = mypch, ylim = ylims, -->
<!-- 		xaxs = 'i') -->
<!--  -->
<!-- TideBoxes(SN05S7$Exposure, SN05S7$DateTimeAST, cols = c('lightblue','white')) -->
<!--  -->
<!-- # Inside pool -->
<!-- loessFunc(SN05S1$DateTimeAST, SN05S1$heartrate, col = 4, pch = mypch,cex=mycex, -->
<!-- 		span = myspan) -->
<!-- loessFunc(SN05S3$DateTimeAST, SN05S3$heartrate, col = 4, pch = mypch,cex=mycex, -->
<!-- 		span = myspan) -->
<!-- loessFunc(SN05S6$DateTimeAST, SN05S6$heartrate, col = 4, pch = mypch,cex=mycex, -->
<!-- 		span = myspan) -->
<!--  -->
<!-- # Outside pool -->
<!-- #loessFunc(SN04S4$DateTimeAST, SN04S4$heartrate, col = 2, pch = mypch, cex = mycex, -->
<!-- #		span = myspan) -->
<!-- loessFunc(SN05S7$DateTimeAST, SN05S7$heartrate, col = 2, pch = mypch, cex = mycex, -->
<!-- 		span = myspan) -->
<!-- # loessFunc(SN04S8$DateTimeAST, SN04S8$heartrate, col = 3, pch = mypch, cex = mycex) -->
<!-- #loessFunc(SN03S4$DateTimeAST, SN03S4$heartrate, col = 1, pch = mypch, cex = mycex, span = myspan) -->
<!--  -->
<!--  -->
<!-- ``` -->






